
\section{Potential Roadblocks \& Alternative Strategies}

\subsection{Insufficient Data}

Issue: Published literature might not contain enough formulations to meet the criteria set above for 30+ data points. Concentration points might also be unevenly distributed, with outliers depending on formulations. 

Alternative Approaches: 
\begin{itemize}
    \item Lower formulations \#: Around $20-25$
    \item Catalyst Types: Expand to more metal additives than iron/magnesium-based
    \item Time Range: Use pre-2000s publications with a hint of scrutiny
\end{itemize}

\subsection{LAMMPS Learning Curve}

Issue: LAMMPS with ReaxFF parameters has a very steep learning curve, which might take more time than anticipated. Simulations could fail to converge, require more powerful computational resources, or unreasonable behavior. 

Alternative Approaches: 
\begin{itemize}
    \item Use smaller simulation cells or fewer atoms for computational demands
    \item Just drop hypothesis 3, as it explored enhancement rather than requirements
    \item Use cloud computing platforms instead of local (Google Colab or AWS)
\end{itemize}

\subsection{Inconsistent Literature}

Issue: Published papers could report data under non-standardized conditions, use different units, or omit formulation details. Uncertainties could be excluded, which would make the analysis less accurate. 

Alternative Approaches: 
\begin{itemize}
    \item Quality Tiers: Categorizing data by reliability and performing analysis while keeping this discrepancy notated within visuals. 
    \item Testing different conclusions depending in the inclusion/exclusion of lower-quality data points.
    \item Use non-parametric tests if data quality is heterogeneous
\end{itemize}

\subsection{Time Constraints}

Issue: The project timeline listed below, and all the goals above, might not be feasible within the 4 months provided. Academic obligations, technical issues, or complications could decrease testing time before due dates. 

Alternative Approaches: 
\begin{itemize}
    \item Prioritize main question over secondaries; treat LAMMPS as an optional enhancement
    \item Reduce formulation count if necessary
    \item Use established software packages (scipy, R) rather than developing custom code
\end{itemize}

